---
title: "Final_Report"
author: "Brisa Banuelos and Quynh Pham"
date: "2026-04-19"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---
```{r setup, include=FALSE}
library(tidyverse)

nhis <- read.csv("data/nhis_clean.csv")
```

# Introduction
This report goes through a variety of health variables and shows how they could be connected. These variables include quantitative variables, such as, age, weight, and height. Qualitative variables include sex of sample adult, race/ethnicity, educational level, general health status, and life satisfaction. 

# Methods
After loading the data into R studio, the data was cleaned and checked for missing values in the variables. After cleaning the data, there were multiple methods used to display the connection between certain variables. Some of these methods included showing a summary of the statistics for the quantitative variables, creating frequency tables for qualitative variables, creating bar plots, box plots, and scatter plots to compare the two types of variables. 

# Results
To begin, we first have to visualize the distribution of the health variables of the participants. When looking at age distribution through a histogram, there was a somewhat even distribution of the participant's ages. The more "popular" ages of the participants included around 63, 71, and 55 years old. The weight distribution of the participants showed a histogram slightly skewed to the right. Some of the "popular" weights included around 180, 160, and 149 pounds. The height distribution appears approximately normal, with no strong skew. A few of the "popular" heights included being around 64, 66, and 67 inches. Lastly, the data appeared to have slightly more female participants than male participants. The scatter plots that compare height vs weight show that as the height of a participant increases, so does the participant's weight.
```{r}
cor(nhis$HEIGHTTC_A, nhis$WEIGHTLBTC_A)
```
The correlation between height and weight is positive, indicating a linear relationship where taller individuals tend to weigh more.
When looking at the enhanced scatter plot that compares height and weight by sex and educational level, it shows that as height increases, weight also increases, and that the majority of male participants are heavier and taller than the female participants. Additionally, a large proportion of participants fall into the high school graduate level and college graduate or better level. Lastly, the correlation matrix showed a weaker or slightly negative relationship between age and weight, a negative relationship between age and height, and a positive relationship between weight and height. 

# Discussion
After reviewing the results from the data collected and visuals created, there is a clear relationship between weight and height, which is, as a participant's height increases, their weight increases as well. Additionally, males were more likely to have a higher weight and height compared to the female participants, as well a large proportion of participants either being high school graduates or college graduates. The findings suggest that demographic factors such as sex and education level are associated with differences in height and weight, although further analysis would be needed to determine causation.

# Conclusion
By reviewing the data collected, the associations between certain health variables, such as weight, height, sex, and educational level, are able to be seen. For example, there was a positive relationship between height and weight consistently throughout the different visualizations created, such as scatterplots and the correlation matrix. 

