---
title: "r markdown gp2"
author: "Daniel Gomez"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
# Day 1: Data Import
#Load these libraries
library(tidyverse)

#Import dataset, keep the csv in the data folder
nhis_data <- read.csv("data/NHIS _Data_2021.csv")
#Initial Exploration no cleaning
cat("Data Structure:\n")
str(nhis_data)
cat("\nSumarry Statistics:\n")
summary(nhis_data)

cat("\nFirst 6 Rows:\n")
head(nhis_data)
```

```{r}
# Day 2: Data Cleaning
library(tidyverse)

#Need to pick a variable and check if anythings missing
nhis_clean <- nhis_data %>%
# Selecting only what I need
select(AGEP_A, WEIGHTLBTC_A, HEIGHTTC_A, SEX_A, HISPALLP_A, 
         EDUCP_A, PHSTAT_A, LSATIS4R_A) %>%
#Remove missing values, look at the codebook
filter(
    !AGEP_A %in% c(97, 98, 99),
    !WEIGHTLBTC_A %in% c(996, 997, 998, 999),
    !HEIGHTTC_A %in% c(96, 97, 98, 99),
    !SEX_A %in% c(7, 9),
    !EDUCP_A %in% c(97, 98, 99),
    !PHSTAT_A %in% c(7, 8, 9),
    !LSATIS4R_A %in% c(7, 8, 9)
  )
# got to use the recode command
nhis_clean <- nhis_clean %>%
  mutate(
#Recode EDUCP_A
    EDUCP_A_recoded = case_when(
      EDUCP_A %in% 0:4 ~ "1 - Less than High School",
      EDUCP_A == 4 ~ "2 - High School Graduate", # (Note: adjusted to fit instructions, usually 0-3 and 4)
      EDUCP_A %in% 5:7 ~ "3 - Some College",
      EDUCP_A %in% 8:10 ~ "4 - College Grad or higher",
      TRUE ~ NA_character_
    ),
#Convert to factor
EDUCP_A_recoded = factor(EDUCP_A_recoded, 
                             levels = c("1 - Less than High School", 
                                        "2 - High School Graduate", 
                                        "3 - Some College", 
                                        "4 - College Grad or higher")),
#Format the other Qualitative stuff to make it easier to read
SEX_A = factor(SEX_A, levels = c(1, 2), labels = c("Male", "Female")),
    PHSTAT_A = factor(PHSTAT_A, levels = c(1,2,3,4,5), 
                      labels = c("Excellent", "Very Good", "Good", "Fair", "Poor")),
    LSATIS4R_A = factor(LSATIS4R_A, levels = c(1,2,3,4), 
                        labels = c("Very Satisfied", "Satisfied", "Dissatisfied", "Very Dissatisfied"))
  ) %>%
#Drop any rows that had NA
drop_na()

#This is to save it
write_csv(nhis_clean, "data/nhis_clean.csv")
print("Data cleaning complete. File saved to data/nhis_clean.csv")
```
```{r}
#Day 3: The Univariate Information
library(tidyverse)

nhis_clean <- read_csv("data/nhis_clean.csv") %>%
  mutate(across(c(SEX_A, EDUCP_A_recoded, PHSTAT_A, LSATIS4R_A), as.factor))
# Quantitative Var
quant_vars <-c("AGEP_A", "WEIGHTLBTC_A", "HEIGHTTC_A")

#Summary I need to see
summary(nhis_clean[quant_vars])

#Standard Deviation
sapply(nhis_clean[quant_vars], sd)
# Somebody fill the rest on line 90 onward



```
```{r}
# Histograms
print(
ggplot(nhis_clean, aes(x = AGEP_A)) +
  geom_histogram() +
  labs(title = "Age Distribution", x = "Age", y = "Count") +
  theme_minimal()
)

print(
ggplot(nhis_clean, aes(x = WEIGHTLBTC_A)) +
  geom_histogram() +
  labs(title = "Weight Distribution", x = "Weight", y = "Count") +
  theme_minimal()
)

print(
ggplot(nhis_clean, aes(x = HEIGHTTC_A)) +
  geom_histogram() +
  labs(title = "Height Distribution", x = "Height", y = "Count") +
  theme_minimal()
)
```

```{r}
# Boxplots
print(
ggplot(nhis_clean, aes(y = AGEP_A)) +
  geom_boxplot() +
  labs(title = "Age Boxplot", y = "Age") +
  theme_minimal()
)

print(
ggplot(nhis_clean, aes(y = WEIGHTLBTC_A)) +
  geom_boxplot() +
  labs(title = "Weight Boxplot", y = "Weight") +
  theme_minimal()
)

print(
ggplot(nhis_clean, aes(y = HEIGHTTC_A)) +
  geom_boxplot() +
  labs(title = "Height Boxplot", y = "Height") +
  theme_minimal()
)
```

```{r}
# Bar Plots
print(
ggplot(nhis_clean, aes(x = SEX_A)) +
  geom_bar() +
  labs(title = "Sex Distribution", x = "Sex", y = "Count") +
  theme_minimal()
)

print(
ggplot(nhis_clean, aes(x = EDUCP_A_recoded)) +
  geom_bar() +
  labs(title = "Education Distribution", x = "Education", y = "Count") +
  theme_minimal()
)

```

# if someone can please check lines 13, 75, 82, that would be great. Whenever I try to run it, it can't find the file. Please and thank you
#Do new project,version control,git,paste the url, type a letter onto the r markdown hit save it should give the option to save the project file to a spot of your choosing on your pc then stage your files in the top right git pane and hit commit leave a message and then push, remember where this project file is stored and open that and do pull from that file, also run the tasks in r markdown in order since there is a clean csv we are working from in day 2
